---
title: "Data exploration for Cm Howicks 2019"
output: html_notebook
---

This document summarizes the initial data exploration for capture data of green turtles at Howick's group of islands in the northern Great Barrier Reef. Data are from Ian Bell. 

```{r}
library(tidyverse)
library(readr)
library(reshape)
library(RMark)
library(lubridate)

save.file <- F

col.def <- cols(PTAG = col_character(),
                TAG = col_integer(),
                TAGST = col_character(),
                SP = col_character(),
                SEX = col_character(),
                AGECL = col_character(),
                CDATE = col_date(format = "%m/%d/%Y"),
                LOCAL = col_character(),
                YEAR = col_integer(),
                HABIT = col_character(),
                PRIAC = col_character(),
                SECAC = col_character(),
                TERAC = col_character(),
                TERAC2 = col_character(),
                EXPER1 = col_character(),
                EXPER2 = col_character(),
                EXPER3 = col_character(),
                TCLUT = col_character(),
                CCL = col_double())

dat.1 <- read_csv(file = "data/juve_Cm_v1.csv", col_types = col.def)

dat.1 %>% mutate(PTAG = as.factor(PTAG),
                 TAGST = as.factor(TAGST),
                 SP = as.factor(SP),
                 SEX = as.factor(SEX),
                 AGECL = as.factor(AGECL),
                 LOCAL = as.factor(LOCAL),
                 HABIT = as.factor(HABIT)) %>%
  transmute(ID = paste(PTAG, TAG, sep = "-"),
            detect = 1,
            TAGST = TAGST,
            SEX = SEX,
            DATE = CDATE,
            LOCAL = LOCAL,
            YEAR = year(CDATE),
            CCL = CCL)-> dat.1
summary(dat.1)
```

Create mark input data file.  Here is a good source for this process (https://sites.google.com/site/cmrsoftware/lecture-lab-schedule/3-intro-to-mark-rmark/formatting-data-for-mark-and-rmark):

```{r}

years <- data.frame(YEAR = min(dat.1$YEAR): max(dat.1$YEAR))

tmp <-melt(dat.1, 
           id.var = c("ID", "YEAR"), 
           measure.var = "detect")

dat.01 <- cast(tmp, ID ~ YEAR)
dat.01$'2000' <- 0
dat.01$'2001' <- 0
dat.01$'2002' <- 0
dat.01$'2003' <- 0
dat.01$'2009' <- 0
dat.01$'2010' <- 0
dat.01$'2011' <- 0
dat.01$'2012' <- 0
dat.01$'2013' <- 0
dat.01$'2019' <- 0

dat.01.1 <- dat.01[, 2:ncol(dat.01)]
dat.01.1[dat.01.1 > 1] <- 1
dat.01 <- cbind(dat.01$ID, 
                dat.01.1[order(names(dat.01.1))]) %>%
  mutate(ID = dat.01$ID) %>%
  select(-"dat.01$ID")

# from https://www.r-bloggers.com/r-recipe-reordering-columns-in-a-flexible-way/
dat.01 <- dat.01[, c("ID", setdiff(names(dat.01), "ID"))]

if (save.file)
  write.csv(dat.01, file = "data/juve_Cm_01_v1.csv", 
            row.names = F,
            quote = F)

c <- ncol(dat.01)
r <- nrow(dat.01)
out.01 <- array(dim = r)
for (k in 1:r){
  out.01[k] <- paste0(dat.01[k, 2:c], collapse = "")
}

caps.hist <- data.frame(ch = out.01, freq = 1)



```

Create covariates
```{r}
# time between two captures:
#cap.dates <- as.Date(colnames(dat.01[2:c]))
#time.dif <- cap.dates[2:c] - cap.dates[1:(c-1)]

# number of capture days per year:
dat.1 %>% select(DATE) -> tmp1
dat.1.date <- unique(tmp1) 

dat.1.date %>%
  mutate(YEAR = year(DATE)) %>%
  group_by(YEAR) %>% 
  summarise(days = n()) %>%
  right_join(years, by = "YEAR") %>%
  transmute(time = as.factor(YEAR),
            days = days) -> catch.days

catch.days[is.na(catch.days$days), "days"] <- 0
```

Try running simple CJS models. I totally forget how these things work... why would phi for non-sampling years come out to be 1.0 and capture probability non zero, when using # days per year as a covariate? 

```{r}
dp <- process.data(caps.hist, 
                   begin.time = 1996, 
                   model = "CJS")

ddl <- make.design.data(dp)
ddl$p <- merge_design.covariates(ddl$p, catch.days)

# models:
dot <- list(formula = ~1)
time <- list(formula = ~time)
effort <- list(formula = ~days)

# create formulas for Phi
Phi.dot <- dot
Phi.time <- time 

#create formulas for p
p.effort <- effort

model.pEffort.phidot <- mark(dp,ddl,
                             model.parameters=list(p=p.effort,
                                                   Phi=Phi.dot), 
                             output=FALSE,
                             threads = 4)

model.pEffort.phiT <- mark(dp,ddl,
                           model.parameters=list(p=p.effort,
                                                 Phi=Phi.time), 
                           output=FALSE,
                           threads = 4)
```

