---
title: "Estimating juvenile survival rates of green turtles at Howicks group of islands"
output: html_notebook
---

This document describes modeling of juvenile survival rates and capture probabilities through Cormack-Jolly-Seber model and the Bayesian approach. The model is from Kery and Schaub. It uses CCL as a covariate. 

```{r}
rm(list=ls())
library(jagsUI)
library(tidyverse)
library(lubridate)
library(reshape)
library(bayesplot)
library(ggridges)


source("Cm_Howicks_functions.R")

# initial values - when removing z, no input
initsFunction <- function(ch){
  phi <- rbeta(1,2,2)
  p <- rbeta(1,2,2)
  z <- known.state.cjs(ch)
  b <- rnorm(1, 0, 50)
  A <- list(mean.phi = phi, mean.p = p, z = z, beta = b)
  return(A)
}

MCMC.params <- list(model.file = "models/Model_CJS_pt.txt",
                       n.chains = 3,
                       n.samples = 50000,
                       n.burnin = 30000,
                       n.thin = 5)

```

Get data - from Ian Bell.

```{r}
col.def <- cols(PTAG = col_character(),
                TAG = col_integer(),
                TAGST = col_character(),
                SP = col_character(),
                SEX = col_character(),
                AGECL = col_character(),
                CDATE = col_date(format = "%m/%d/%Y"),
                LOCAL = col_character(),
                YEAR = col_integer(),
                HABIT = col_character(),
                PRIAC = col_character(),
                SECAC = col_character(),
                TERAC = col_character(),
                TERAC2 = col_character(),
                EXPER1 = col_character(),
                EXPER2 = col_character(),
                EXPER3 = col_character(),
                TCLUT = col_character(),
                CCL = col_double())

dat.1 <- read_csv(file = "data/juve_Cm_v1.csv", col_types = col.def)

dat.1 %>% mutate(PTAG = as.factor(PTAG),
                 TAGST = as.factor(TAGST),
                 SP = as.factor(SP),
                 SEX = as.factor(SEX),
                 AGECL = as.factor(AGECL),
                 LOCAL = as.factor(LOCAL),
                 HABIT = as.factor(HABIT)) %>%
  transmute(ID = paste(PTAG, TAG, sep = "-"),
            detect = 1,
            TAGST = TAGST,
            SEX = SEX,
            DATE = CDATE,
            LOCAL = LOCAL,
            HABIT = HABIT,
            YEAR = YEAR,
            CCL = CCL)-> dat.1
```

Set up data for analysis.  Captures within each season are condensed to one column.  Also, CCL should be extracted for the first capture. 

```{r}
CJS.data <- dat2CJS(dat.1, save.file = FALSE)
CJS.data$data %>% rownames_to_column(var = "ID") -> data.CJS

```

Then get ready for CJS anlaysis

```{r}
# capture history
CH <- as.matrix(CJS.data$data)

nInd <- nrow(CJS.data$data)

ns <- colSums(CJS.data$data)

# find the first capture date
get.first <- function(x) min(which(x != 0))
f <- apply(CH, 1, get.first)

## parameters to monitor - when this is changed, make sure to change
## summary statistics indices at the end of this script. 
parameters <- c("p", "mean.phi", "N")

jags.data <- list(y = CH, f = f, nind = nInd, 
                  n.occasions = dim(CH)[2],
                  n = ns, T = ncol(CJS.data$data))

inits <- lapply(replicate(MCMC.params$n.chains, 
                          CH, 
                          simplify = FALSE), 
                initsFunction)

jm <- jags(data = jags.data,
           inits = inits,
           parameters.to.save= parameters,
           model.file = MCMC.params$model.file,
           n.chains = MCMC.params$n.chains,
           n.burnin = MCMC.params$n.burnin,
           n.thin = MCMC.params$n.thin,
           n.iter = MCMC.params$n.samples,
           DIC = T, 
           parallel=T)

out.list <- list(jags.out = jm,
                   jags.data = jags.data)

 saveRDS(out.list, file = "RData/CJS_pt.rds")

```

```{r}
# look at trace and posteriors
mcmc_trace(jm$samples, c("mean.phi"))
mcmc_dens(jm$samples, c("mean.phi"))
```

```{r}
# look at trace and posteriors
mcmc_dens(jm$samples, c("p[1]", "p[2]", "p[3]", "p[4]", "p[5]", "p[6]"))
mcmc_dens(jm$samples, c("p[7]","p[8]", "p[9]", "p[10]", "p[11]", "p[12]", "p[13]"))
#mcmc_dens(jm$samples, c("mean.phi"))

```

```{r}
# look at trace and posteriors
mcmc_dens(jm$samples, c("N[1]", "N[2]", "N[3]", "N[4]", "N[5]", "N[6]"))
mcmc_dens(jm$samples, c("N[7]","N[8]", "N[9]", "N[10]", "N[11]", "N[12]", "N[13]"))
#mcmc_dens(jm$samples, c("mean.phi"))

```

```{r}
Ns.df <- extract.posterior("N", jm$samples)
colnames(Ns.df) <- sort(unique(dat.1$YEAR))

Ns.df.2 <- melt(Ns.df, 
                value.name = "N", 
                variable.name = "Year")

colnames(Ns.df.2) <- c("Year", "N")

ggplot(data = Ns.df.2) + 
  geom_density_ridges2(aes(y = Year, x = N),
                       fill = "red") + 
  xlim(c(0, 2500))
```

