---
title: "Annual abundance estimates for juvenile green turtles at Ingram Island"
output: html_notebook
---

This document describes esimating abundance of juvenile green turtles at Ingram Island from capture-recapture data.  

Set up some constants and load libraries and define functions. 

```{r}
rm(list=ls())
library(jagsUI)
library(tidyverse)
library(lubridate)
library(reshape)
library(bayesplot)

source("Cm_Howicks_functions.R")

SAVE <- TRUE

MCMC.params.M0 <- list(model.file = "models/Model_M0.txt",
                       n.chains = 3,
                       n.samples = 50000,
                       n.burnin = 30000,
                       n.thin = 5)

MCMC.params.Mt <- list(model.file = "models/Model_Mt.txt",
                       n.chains = 3,
                       n.samples = 50000,
                       n.burnin = 30000,
                       n.thin = 5)

```


First, load capture history data from Ian Bell. 

```{r}
col.def <- cols(PTAG = col_character(),
                TAG = col_integer(),
                TAGST = col_character(),
                SP = col_character(),
                SEX = col_character(),
                AGECL = col_character(),
                CDATE = col_date(format = "%m/%d/%Y"),
                LOCAL = col_character(),
                YEAR = col_integer(),
                HABIT = col_character(),
                PRIAC = col_character(),
                SECAC = col_character(),
                TERAC = col_character(),
                TERAC2 = col_character(),
                EXPER1 = col_character(),
                EXPER2 = col_character(),
                EXPER3 = col_character(),
                TCLUT = col_character(),
                CCL = col_double())

dat.1 <- read_csv(file = "data/juve_Cm_v3_2019.csv", 
                  col_types = col.def)

dat.1 %>% mutate(PTAG = as.factor(PTAG),
                 TAGST = as.factor(TAGST),
                 SP = as.factor(SP),
                 SEX = as.factor(SEX),
                 AGECL = as.factor(AGECL),
                 LOCAL = as.factor(LOCAL),
                 HABIT = as.factor(HABIT)) %>%
  transmute(ID = paste(PTAG, TAG, sep = "-"),
            detect = 1,
            TAGST = TAGST,
            SEX = SEX,
            DATE = CDATE,
            LOCAL = LOCAL,
            HABIT = HABIT,
            YEAR = year(CDATE),
            CCL = CCL)-> dat.1

#summary(dat.1)
```

Data are from 1997 to 2018. Just to get things started, I'll start from 2014 when we had a lot of recaptures. 

```{r}
# filter just one year
year <- 2014
out <- dat2dat01(dat.1, YEAR = year, save.file = TRUE)

if (SAVE == TRUE) saveFname_M0 <- paste("RData/M0_", year, "_",
                                     Sys.Date(), ".rds", 
                                     sep = "")

if (SAVE == TRUE) saveFname_Mt <- paste("RData/Mt_", year, "_",
                                     Sys.Date(), ".rds", 
                                     sep = "")

data01 <- out$data[1:nrow(out$data), 
                   2:ncol(out$data)]

col.sums <- colSums(data01)
data01 <- data01[, col.sums > 20]

# remove rows with all zeros
row.sums <- rowSums(data01)
data01 <- data01[row.sums != 0,]

if (!file.exists(saveFname_M0)){
  
  jm.M0 <- jags.M0(data01, 
                   nz = 500, 
                   MCMC.params = MCMC.params.M0)

  out.list.M0 <- list(jags.out = jm.M0,
                      MCMC.params = MCMC.params.M0,
                      model.file = MCMC.params.M0$model.file)
  
  if (SAVE == TRUE) {
    saveRDS(out.list.M0, file = saveFname_M0)
  }  
}

if (!file.exists(saveFname_Mt)){
  
  jm.Mt <- jags.Mt(data01, 
                   nz = 500, 
                   MCMC.params = MCMC.params.Mt)
  
  
  out.list.Mt <- list(jags.out = jm.Mt,
                      MCMC.params = MCMC.params.Mt,
                      model.file = MCMC.params.Mt$model.file)
  
  if (SAVE == TRUE) {
    saveRDS(out.list.Mt, file = saveFname_Mt)
  }  
  
}

```


```{r}
# look at trace and posteriors
mcmc_trace(jm.M0$jags.out$samples, c("N", "p"))

mcmc_dens(jm.M0$jags.out$samples, c("N", "p"))

```


```{r}
# look at trace and posteriors
mcmc_trace(jm.Mt$jags.out$samples, 
           c("N", "p[1]", "p[2]", "p[3]", "p[4]", "p[5]"))

mcmc_dens(jm.Mt$jags.out$samples, 
          c("N", "p[1]", "p[2]", "p[3]", "p[4]", "p[5]"))

```

```{r}
# filter just one year
year <- 2017
out <- dat2dat01(dat.1, YEAR = year, save.file = TRUE)

if (SAVE == TRUE) saveFname_M0 <- paste("RData/M0_", year, "_",
                                     Sys.Date(), ".rds", 
                                     sep = "")

if (SAVE == TRUE) saveFname_Mt <- paste("RData/Mt_", year, "_",
                                     Sys.Date(), ".rds", 
                                     sep = "")

data01 <- out$data[1:nrow(out$data), 
                   2:ncol(out$data)]

col.sums <- colSums(data01)
data01 <- data01[, col.sums > 10]

# remove rows with all zeros
row.sums <- rowSums(data01)
data01 <- data01[row.sums != 0,]

if (!file.exists(saveFname_M0)){
  
  jm.M0 <- jags.M0(data01, 
                   nz = 500, 
                   MCMC.params = MCMC.params.M0)
  
  out.list.M0 <- list(jags.out = jm.M0,
                      MCMC.params = MCMC.params.M0,
                      model.file = MCMC.params.M0$model.file)

  if (SAVE == TRUE) {
      saveRDS(out.list.M0, file = saveFname_M0)
  }  
}

if (!file.exists(saveFname_Mt)){

  jm.Mt <- jags.Mt(data01, 
                   nz = 500, 
                   MCMC.params = MCMC.params.Mt)
  
  
  out.list.Mt <- list(jags.out = jm.Mt,
                   MCMC.params = MCMC.params.Mt,
                   model.file = MCMC.params.Mt$model.file)

  if (SAVE == TRUE) {
      saveRDS(out.list.Mt, file = saveFname_Mt)
  }  

}

```


```{r}
# look at trace and posteriors
mcmc_trace(jm.Mt$jags.out$samples, c("N", "p[1]", "p[2]", "p[3]"))

mcmc_dens(jm.Mt$jags.out$samples, c("N", "p[1]", "p[2]", "p[3]"))

```

```{r}
# filter just one year
year <- 2018

out <- dat2dat01(dat.1, year = year, save.file = TRUE)

if (SAVE == TRUE) saveFname_M0 <- paste("RData/M0_", year, "_",
                                     Sys.Date(), ".rds", 
                                     sep = "")

if (SAVE == TRUE) saveFname_Mt <- paste("RData/Mt_", year, "_",
                                     Sys.Date(), ".rds", 
                                     sep = "")

data01 <- out$data[1:nrow(out$data), 
                   2:ncol(out$data)]

col.sums <- colSums(data01)
data01 <- data01[, col.sums > 9]

# remove rows with all zeros
row.sums <- rowSums(data01)
data01 <- data01[row.sums != 0,]

if (!file.exists(saveFname_M0)){
  
  jm.M0 <- jags.M0(data01, 
                   nz = 500, 
                   MCMC.params = MCMC.params.M0)
  
  out.list.M0 <- list(jags.out = jm.M0,
                      MCMC.params = MCMC.params.M0,
                      model.file = MCMC.params.M0$model.file)

  if (SAVE == TRUE) {
      saveRDS(out.list.M0, file = saveFname_M0)
  }  
}

if (!file.exists(saveFname_Mt)){

  jm.Mt <- jags.Mt(data01, 
                   nz = 500, 
                   MCMC.params = MCMC.params.Mt)
  
  
  out.list.Mt <- list(jags.out = jm.Mt,
                   MCMC.params = MCMC.params.Mt,
                   model.file = MCMC.params.Mt$model.file)

  if (SAVE == TRUE) {
      saveRDS(out.list.Mt, file = saveFname_Mt)
  }  
}
```


```{r}
# look at trace and posteriors
mcmc_trace(jm.Mt$jags.out$samples, c("N", "p[1]", "p[2]", "p[3]", "p[4]", "p[5]"))

mcmc_dens(jm.Mt$jags.out$samples, c("N", "p[1]", "p[2]", "p[3]", "p[4]", "p[5]"))


```


```{r}
# filter just one year
year <- 2004

out <- dat2dat01(dat.1, YEAR = year, save.file = TRUE)

if (SAVE == TRUE) saveFname_M0 <- paste("RData/M0_", year, "_",
                                     Sys.Date(), ".rds", 
                                     sep = "")

if (SAVE == TRUE) saveFname_Mt <- paste("RData/Mt_", year, "_",
                                     Sys.Date(), ".rds", 
                                     sep = "")

data01 <- out$data[1:nrow(out$data), 
                   2:ncol(out$data)]

col.sums <- colSums(data01)
data01 <- data01[, col.sums > 10]

# remove rows with all zeros
row.sums <- rowSums(data01)
data01 <- data01[row.sums != 0,]

if (!file.exists(saveFname_M0)){
  
  jm.M0 <- jags.M0(data01, 
                   nz = 500, 
                   MCMC.params = MCMC.params.M0)
  
  out.list.M0 <- list(jags.out = jm.M0,
                      MCMC.params = MCMC.params.M0,
                      model.file = MCMC.params.M0$model.file)

  if (SAVE == TRUE) {
      saveRDS(out.list.M0, file = saveFname_M0)
  }  
}

if (!file.exists(saveFname_Mt)){

  jm.Mt <- jags.Mt(data01, 
                   nz = 500, 
                   MCMC.params = MCMC.params.Mt)
  
  
  out.list.Mt <- list(jags.out = jm.Mt,
                   MCMC.params = MCMC.params.Mt,
                   model.file = MCMC.params.Mt$model.file)

  if (SAVE == TRUE) {
      saveRDS(out.list.Mt, file = saveFname_Mt)
  }  
}
```

```{r}
# look at trace and posteriors
mcmc_trace(jm.Mt$jags.out$samples, c("N", "p[1]", "p[2]"))

mcmc_dens(jm.Mt$jags.out$samples, c("N", "p[1]", "p[2]"))


```

```{r}
# filter just one year
year <- 2016
out <- dat2dat01(dat.1, YEAR = year, save.file = TRUE)

if (SAVE == TRUE) saveFname_M0 <- paste("RData/M0_", year, "_",
                                     Sys.Date(), ".rds", 
                                     sep = "")

if (SAVE == TRUE) saveFname_Mt <- paste("RData/Mt_", year, "_",
                                     Sys.Date(), ".rds", 
                                     sep = "")

data01 <- out$data[1:nrow(out$data), 
                   2:ncol(out$data)]

# there was one ID caught twice in a day QA-51289 on 2916-07-29 so
# need to fix that 

data01[data01 > 1] <- 1

col.sums <- colSums(data01)
data01 <- data01[, col.sums > 18]

# remove rows with all zeros
row.sums <- rowSums(data01)
data01 <- data01[row.sums != 0,]

if (!file.exists(saveFname_M0)){
  
  jm.M0 <- jags.M0(data01, 
                   nz = 500, 
                   MCMC.params = MCMC.params.M0)

  out.list.M0 <- list(jags.out = jm.M0,
                      MCMC.params = MCMC.params.M0,
                      model.file = MCMC.params.M0$model.file)
  
  if (SAVE == TRUE) {
    saveRDS(out.list.M0, file = saveFname_M0)
  }  
}

if (!file.exists(saveFname_Mt)){
  
  jm.Mt <- jags.Mt(data01, 
                   nz = 500, 
                   MCMC.params = MCMC.params.Mt)
  
  
  out.list.Mt <- list(jags.out = jm.Mt,
                      MCMC.params = MCMC.params.Mt,
                      model.file = MCMC.params.Mt$model.file)
  
  if (SAVE == TRUE) {
    saveRDS(out.list.Mt, file = saveFname_Mt)
  }  
  
}

```


```{r}
# look at trace and posteriors
mcmc_trace(jm.Mt$jags.out$samples, c("N", "p[1]", "p[2]", "p[3]", "p[4]", "p[5]"))

mcmc_dens(jm.Mt$jags.out$samples, c("N", "p[1]", "p[2]", "p[3]", "p[4]", "p[5]"))

```

For above years, N seems just way too high... this probably comes from the low numbers of recaptures.  Take a look at how many were recaptured each year. 

```{r}
# recaptures
uniq.yrs <- sort(unique(dat.1$YEAR))
  #1997, 1998, 1999 have so few data, skip:
uniq.yrs <- uniq.yrs[uniq.yrs != 1997]
uniq.yrs <- uniq.yrs[uniq.yrs != 1998]
uniq.yrs <- uniq.yrs[uniq.yrs != 1999]

# minimum column sum for each year needs to be done manually... Starts from 2004
min.col.sum <- c(10, 10, 10, 10, 10, 10, 18, 19, 10, 10)

n.captures.list <- list()

k <- 13
for (k in 1:length(uniq.yrs)){
  out <- dat2dat01(dat.1, year = uniq.yrs[k], save.file = FALSE)
  data01 <- out$data[1:nrow(out$data), 
                     2:ncol(out$data)]
  
  # there was one ID caught twice in a day QA-51289 on 2916-07-29 so
  # need to fix that 
  data01[data01 > 1] <- 1
  
  col.sums <- colSums(data01)

  data01 <- data01[, col.sums > min.col.sum[k]]
# remove rows with all zeros
  row.sums <- rowSums(data01)
  data01 <- data01[row.sums != 0,]
  
  n.captures <- rowSums(data01)
  n.captures.list[[k]] <- table(n.captures)
  
}
```

