---
title: "Estimating juvenile survival rates of green turtles at Howicks group of islands"
output: html_notebook
---

This document describes modeling of juvenile survival rates and capture probabilities through Cormack-Jolly-Seber model and the Bayesian approach. The model is from Kery and Schaub. It uses CCL as a covariate. 

```{r}
rm(list=ls())
library(jagsUI)
library(tidyverse)
library(lubridate)
library(reshape)
library(bayesplot)

dat2CJS <- function(dat.1, save.file = TRUE){
  
  # Create ID by Date and assign 1s
  tmp <-melt(dat.1, 
             id.var = c("ID", "YEAR"), 
             measure.var = "detect")
  
  # make a table with ID by year
  dat.01 <- cast(tmp, ID ~ YEAR)
  
  # replace > 1 with ones
  dat.01 <- as.data.frame(dat.01) %>%
    remove_rownames() %>%
    column_to_rownames(var = "ID")
  
  dat.01[(dat.01 > 1)] <- 1
  
  # save file for later
  if (save.file){
    out.name <- "data/juve_Cm_01_CJS.csv"
    write.csv(dat.01, 
              file = out.name, 
              row.names = T,
              quote = F)
    
  } else {
    out.name <- "not.saved"
  }
  
  out <- list(filename = out.name,
              data = dat.01)
  
  return(out)
              
}

known.state.cjs <- function(ch){
  state <- ch
  for (i in 1:dim(ch)[1]){
    n1 <- min(which(ch[i,]==1))
    n2 <- max(which(ch[i,]==1))
    state[i,n1:n2] <- 1
    state[i,n1] <- NA
  }
  state[state==0] <- NA
  return(state)
}

# initial values - when removing z, no input
initsFunction <- function(ch){
  phi <- rbeta(1,2,2)
  p <- rbeta(1,2,2)
  z <- known.state.cjs(ch)
  b <- rnorm(1, 0, 50)
  A <- list(mean.phi = phi, mean.p = p, z = z, beta = b)
  return(A)
}


MCMC.params <- list(model.file = "models/Model_CJS_firstCCL.txt",
                       n.chains = 3,
                       n.samples = 50000,
                       n.burnin = 30000,
                       n.thin = 5)


```

Get data - from Ian Bell.

```{r}
col.def <- cols(PTAG = col_character(),
                TAG = col_integer(),
                TAGST = col_character(),
                SP = col_character(),
                SEX = col_character(),
                AGECL = col_character(),
                CDATE = col_date(format = "%m/%d/%Y"),
                LOCAL = col_character(),
                YEAR = col_integer(),
                HABIT = col_character(),
                PRIAC = col_character(),
                SECAC = col_character(),
                TERAC = col_character(),
                TERAC2 = col_character(),
                EXPER1 = col_character(),
                EXPER2 = col_character(),
                EXPER3 = col_character(),
                TCLUT = col_character(),
                CCL = col_double())

dat.1 <- read_csv(file = "data/juve_Cm_v1.csv", col_types = col.def)

dat.1 %>% mutate(PTAG = as.factor(PTAG),
                 TAGST = as.factor(TAGST),
                 SP = as.factor(SP),
                 SEX = as.factor(SEX),
                 AGECL = as.factor(AGECL),
                 LOCAL = as.factor(LOCAL),
                 HABIT = as.factor(HABIT)) %>%
  transmute(ID = paste(PTAG, TAG, sep = "-"),
            detect = 1,
            TAGST = TAGST,
            SEX = SEX,
            DATE = CDATE,
            LOCAL = LOCAL,
            HABIT = HABIT,
            YEAR = YEAR,
            CCL = CCL)-> dat.1
```

Set up data for analysis.  Captures within each season are condensed to one column.  Also, CCL should be extracted for the first capture. 

```{r}
CJS.data <- dat2CJS(dat.1, save.file = TRUE)
CJS.data$data %>% rownames_to_column(var = "ID") -> data.CJS

CCL.data <- dat2CCL(dat.1, save.file = TRUE)

data.CJS %>% left_join(CCL.data$data, by = "ID") -> CJS.CCL.data

```

Then get ready for CJS anlaysis

```{r}
# capture history
CH <- as.matrix(CJS.CCL.data[, c("1988", "1997", "1999", "2004", "2005",
                                 "2006", "2007", "2008", "2014", "2015",
                                 "2016", "2017", "2018")])
nInd <- nrow(CH)

CCL_min <- as.matrix(CJS.CCL.data[,"CCL_min"])
CCL_max <- as.matrix(CJS.CCL.data[,"CCL_max"])
CCL_mean <- as.matrix(CJS.CCL.data[,"CCL_mean"])

# find the first capture date
get.first <- function(x) min(which(x != 0))
f <- apply(CH, 1, get.first)

# fill in the CCL for missing values - Use mean of all
CCL_min[is.na(CCL_min)] <- mean(CCL_mean, na.rm = T)
CCL_min[is.infinite(CCL_min)] <- mean(CCL_mean, na.rm = T)

## parameters to monitor - when this is changed, make sure to change
## summary statistics indices at the end of this script. 
parameters <- c("mean.p", "mean.phi", "beta")

jags.data <- list(y = CH, f = f, nind = nInd, 
                  n.occasions = dim(CH)[2],
                  x = as.vector(CCL_min))

inits <- lapply(replicate(MCMC.params$n.chains, 
                          CH, 
                          simplify = FALSE), 
                initsFunction)

jm <- jags(data = jags.data,
           inits = inits,
           parameters.to.save= parameters,
           model.file = MCMC.params$model.file,
           n.chains = MCMC.params$n.chains,
           n.burnin = MCMC.params$n.burnin,
           n.thin = MCMC.params$n.thin,
           n.iter = MCMC.params$n.samples,
           DIC = T, 
           parallel=T)

out.list <- list(jags.out = jm,
                   jags.data = jags.data)

 saveRDS(out.list, file = "RData/CJS_CCL.rds")

```

