---
title: "Juvenile grwoth at Howicks"
output: html_notebook
---

I try to fit von Bertelanffy growth curve to observed growth data of juvenile greens at Howicks. 

I have to convert the raw data files into an input file. The original model was designed for analyzing growths at Milman.  

```{r}

rm(list=ls())
library(jagsUI)
library(tidyverse)
library(lubridate)
library(reshape)
library(bayesplot)
library(ggridges)

save.rds <- TRUE

source("Cm_Howicks_functions.R")

dat.1 <- get.data("data/juve_Cm_v3_2019.csv")

```

Find those that were captured more than once:

```{r}
n.cap.ID <- table(dat.1$ID)
recap.ID <- data.frame(n.cap.ID[n.cap.ID > 1])
colnames(recap.ID) <- c("ID", "Freq")

recap.ID %>% left_join(dat.1, by = "ID") -> recap.data
```

Make length and capture date matrices
```{r}
unique.ID <- recap.ID$ID
size.mat <- date.mat <- matrix(nrow = length(unique.ID),
                              ncol = max(recap.data$Freq))

date.1 <- structure(numeric(length(unique.ID)), class = "Date")
n.vec <- vector(mode = "numeric", length = length(unique.ID))

k <- 1
for (k in 1:length(unique.ID)){
  tmp.ID <- filter(recap.data, ID == unique.ID[k])
  size.mat[k, 1:nrow(tmp.ID)] <- tmp.ID$CCL
  date.mat[k, 1:nrow(tmp.ID)] <- tmp.ID$DATE - min(tmp.ID$DATE)
  date.1[k] <- min(tmp.ID$DATE)
  n.vec[k] <- nrow(tmp.ID)
}

date.mat <- date.mat[, 2:ncol(date.mat)]/365
```

Then, send this to jags with a couple models.

```{r}
MCMC.params <- list(model.file = "models/Model_RlinfRk_L.txt",
                    n.chains = 3,
                    n.samples = 100000,
                    n.burnin = 60000,
                    n.thin = 5)

if (!file.exists( "RData/Rlinf_Rk_L.rds")){
  parameters <- c('CV', 'k', 'A', 
                  'Linf', 'LinfMu', 'LinfSD', 
                  'Shape', 'rate', #'kAlpha', 'kBeta', 
                  'deviance')
  
  jags.data <- list(nIndiv = length(unique.ID),
                    n = n.vec,
                    L = size.mat,
                    t = date.mat)
  
  initsFunction <- function(d){
    #kAlpha <- runif(1, 0, 10)
    #kBeta <- runif(1, 0, 10)
    k <- rbeta(length(unique.ID), 1, 1)
    CV <- rbeta(1, 1, 1)
    Shape <- runif(1, 0, 10)
    rate <- runif(1, 0, 10)
    LinfMu <- runif(1, 90, 200)
    LinfTau <- rgamma(1, 1, 1)
    A <- list(k = k, CV = CV,
              Shape = Shape, rate = rate)
    return(A)
  }
  
  # Create a list of length nChains with z in initsFunction
  #inList <- list(CH, f)
  inits <- lapply(c(1:MCMC.params$n.chains), initsFunction)
  
  jm.1 <- jags(data = jags.data,
               #inits = inits,
               parameters.to.save= parameters,
               model.file = MCMC.params$model.file,
               n.chains = MCMC.params$n.chains,
               n.burnin = MCMC.params$n.burnin,
               n.thin = MCMC.params$n.thin,
               n.iter = MCMC.params$n.samples,
               DIC = T, 
               parallel=T)
  
  if(save.rds) saveRDS(jm.1, file = "RData/Rlinf_Rk_L.rds")
  
} else {
  jm.1 <- readRDS(file = "RData/Rlinf_Rk_L.rds")
}

```

look at the posterior;

```{r}
mcmc_trace(jm.1$samples, c("LinfMu"))
mcmc_dens(jm.1$samples, c("LinfMu"))

```

```{r}
mcmc_trace(jm.1$samples, c("CV"))
mcmc_dens(jm.1$samples, c("CV"))

```

Plot what we found
```{r}
Linf <- jm.1$q50$Linf
k <- jm.1$q50$k
A <- jm.1$q50$A
time.vec <- 0:49

time.mat <- L_Exp <- matrix(ncol = length(unique.ID), nrow = length(time.vec))

i <- 1
for (i in 1:length(unique.ID)){
  L_Exp[,i] <-  Linf[i] * (1.0 - exp(-k[i] * (A[i] + time.vec)))
  time.mat[,i] <- A[i] + time.vec
}

colnames(L_Exp) <- unique.ID
L_Exp_long <- melt(L_Exp)
colnames(L_Exp_long) <- c("time1", "ID", "CCL")

colnames(time.mat) <- unique.ID
time_long <- melt(time.mat)
colnames(time_long) <- c("time1", "ID", "Years")

L_Exp_long$Years <- time_long$Years

ggplot(data = L_Exp_long) + 
  geom_path(aes(x = Years, y = CCL, color = ID)) +
  theme(legend.position = "none")
```

The model seems to fit fine, although the Rhat statistic was not all great so I probably have to increase the number of MCMC samples. 

I also want to run fixed k model
```{r}
if (!file.exists( "RData/Rlinf_Fk.rds")){
  MCMC.params$model.file <- "models/Model_Rlinf_Fk.txt"
  parameters <- c('CV', 'k', 'A', 
                  'Linf', 'LinfMu', 'LinfSD', 
                  'Shape', 'rate', #'kAlpha', 'kBeta', 
                  'deviance')
  
  jags.data <- list(nIndiv = length(unique.ID),
                    n = n.vec,
                    L = size.mat,
                    t = date.mat)
  
  jm.2 <- jags(data = jags.data,
               #inits = inits,
               parameters.to.save= parameters,
               model.file = MCMC.params$model.file,
               n.chains = MCMC.params$n.chains,
               n.burnin = MCMC.params$n.burnin,
               n.thin = MCMC.params$n.thin,
               n.iter = MCMC.params$n.samples,
               DIC = T, 
               parallel=T)
  
  if(save.rds) saveRDS(jm.2, file = "RData/Rlinf_Fk.rds")
  
} else {
  jm.2 <- readRDS(file = "RData/Rlinf_Fk.rds")
}
```


```{r}
mcmc_trace(jm.2$samples, c("k"))
mcmc_dens(jm.2$samples, c("k"))

```

These don't look so good... 

Next, try to find growth rates between early years and the last 5 years. 

```{r}
dat.1 %>% filter(DATE < as.Date("2015-01-01")) -> dat.1.pre2015
dat.1 %>% filter(DATE >= as.Date("2015-01-01")) -> dat.1.post2015

n.cap.ID.pre2015 <- table(dat.1.pre2015$ID)
recap.ID.pre2015 <- data.frame(n.cap.ID.pre2015[n.cap.ID.pre2015 > 1])
colnames(recap.ID.pre2015) <- c("ID", "Freq")

recap.ID.pre2015 %>% left_join(dat.1.pre2015, by = "ID") -> recap.data.pre2015

n.cap.ID.post2015 <- table(dat.1.post2015$ID)
recap.ID.post2015 <- data.frame(n.cap.ID.post2015[n.cap.ID.post2015 > 1])
colnames(recap.ID.post2015) <- c("ID", "Freq")

recap.ID.post2015 %>% left_join(dat.1.post2015, by = "ID") -> recap.data.post2015

```
```{r}
unique.ID.pre2015 <- recap.ID.pre2015$ID
size.mat.pre2015 <- date.mat.pre2015 <- matrix(nrow = length(unique.ID.pre2015),
                                               ncol = max(recap.data.pre2015$Freq))

date.1.pre2015 <- structure(numeric(length(unique.ID.pre2015)), class = "Date")
n.vec.pre2015 <- vector(mode = "numeric", length = length(unique.ID.pre2015))

k <- 1
for (k in 1:length(unique.ID.pre2015)){
  tmp.ID <- filter(recap.data.pre2015, ID == unique.ID.pre2015[k])
  size.mat.pre2015[k, 1:nrow(tmp.ID)] <- tmp.ID$CCL
  date.mat.pre2015[k, 1:nrow(tmp.ID)] <- tmp.ID$DATE - min(tmp.ID$DATE)
  date.1.pre2015[k] <- min(tmp.ID$DATE)
  n.vec.pre2015[k] <- nrow(tmp.ID)
}

date.mat.pre2015 <- date.mat.pre2015[, 2:ncol(date.mat.pre2015)]/365

unique.ID.post2015 <- recap.ID.post2015$ID
size.mat.post2015 <- date.mat.post2015 <- matrix(nrow = length(unique.ID.post2015),
                                               ncol = max(recap.data.post2015$Freq))

date.1.post2015 <- structure(numeric(length(unique.ID.post2015)), class = "Date")
n.vec.post2015 <- vector(mode = "numeric", length = length(unique.ID.post2015))

k <- 1
for (k in 1:length(unique.ID.post2015)){
  tmp.ID <- filter(recap.data.post2015, ID == unique.ID.post2015[k])
  size.mat.post2015[k, 1:nrow(tmp.ID)] <- tmp.ID$CCL
  date.mat.post2015[k, 1:nrow(tmp.ID)] <- tmp.ID$DATE - min(tmp.ID$DATE)
  date.1.post2015[k] <- min(tmp.ID$DATE)
  n.vec.post2015[k] <- nrow(tmp.ID)
}

date.mat.post2015 <- date.mat.post2015[, 2:ncol(date.mat.post2015)]/365

```

run the models on both datasets

```{r}
MCMC$model.file <- "models/Model_RlinfRk_L.txt"

if (!file.exists( "RData/Rlinf_Rk_pre2015.rds")){
  parameters <- c('CV', 'k', 'A', 
                  'Linf', 'LinfMu', 'LinfSD', 
                  'Shape', 'rate', #'kAlpha', 'kBeta', 
                  'deviance')
  
  jags.data <- list(nIndiv = length(unique.ID.pre2015),
                    n = n.vec.pre2015,
                    L = size.mat.pre2015,
                    t = date.mat.pre2015)
  
  jm.3 <- jags(data = jags.data,
               #inits = inits,
               parameters.to.save= parameters,
               model.file = MCMC.params$model.file,
               n.chains = MCMC.params$n.chains,
               n.burnin = MCMC.params$n.burnin,
               n.thin = MCMC.params$n.thin,
               n.iter = MCMC.params$n.samples,
               DIC = T, 
               parallel=T)
  
  if(save.rds) saveRDS(jm.3, file = "RData/Rlinf_Rk_pre2015.rds")
  
} else {
  jm.3 <- readRDS(file = "RData/Rlinf_Rk_pre2015.rds")
}

if (!file.exists( "RData/Rlinf_Rk_post2015.rds")){
  parameters <- c('CV', 'k', 'A', 
                  'Linf', 'LinfMu', 'LinfSD', 
                  'Shape', 'rate', #'kAlpha', 'kBeta', 
                  'deviance')
  
  jags.data <- list(nIndiv = length(unique.ID.post2015),
                    n = n.vec.post2015,
          
                    L = size.mat.post2015,
                    t = date.mat.post2015)
  
  jm.4 <- jags(data = jags.data,
               #inits = inits,
               parameters.to.save= parameters,
               model.file = MCMC.params$model.file,
               n.chains = MCMC.params$n.chains,
               n.burnin = MCMC.params$n.burnin,
               n.thin = MCMC.params$n.thin,
               n.iter = MCMC.params$n.samples,
               DIC = T, 
               parallel=T)
  
  if(save.rds) saveRDS(jm.4, file = "RData/Rlinf_Rk_post2015.rds")
  
} else {
  jm.4 <- readRDS(file = "RData/Rlinf_Rk_post2015.rds")
}

```


```{r}
mcmc_trace(jm.3$samples, c("CV"))
mcmc_dens(jm.3$samples, c("CV"))

```


```{r}
Linf <- jm.3$q50$Linf
k <- jm.3$q50$k
A <- jm.3$q50$A
time.vec <- 0:19

time.mat.pre2015 <- L_Exp.pre2015 <- matrix(ncol = length(unique.ID.pre2015), 
                                            nrow = length(time.vec))

i <- 1
for (i in 1:length(unique.ID.pre2015)){
  L_Exp.pre2015[,i] <-  Linf[i] * (1.0 - exp(-k[i] * (A[i] + time.vec)))
  time.mat.pre2015[,i] <- A[i] + time.vec
}

colnames(L_Exp.pre2015) <- unique.ID.pre2015
L_Exp_long.pre2015 <- melt(L_Exp.pre2015)
colnames(L_Exp_long.pre2015) <- c("time1", "ID", "CCL")

colnames(time.mat.pre2015) <- unique.ID.pre2015
time_long.pre2015 <- melt(time.mat.pre2015)
colnames(time_long.pre2015) <- c("time1", "ID", "Years")

L_Exp_long.pre2015$Years <- time_long.pre2015$Years

ggplot(data = L_Exp_long.pre2015) + 
  geom_path(aes(x = Years, y = CCL, color = ID)) +
  theme(legend.position = "none")
```



```{r}
Linf <- jm.4$q50$Linf
k <- jm.4$q50$k
A <- jm.4$q50$A
time.vec <- 0:60

time.mat.post2015 <- L_Exp.post2015 <- matrix(ncol = length(unique.ID.post2015), 
                                            nrow = length(time.vec))

i <- 1
for (i in 1:length(unique.ID.post2015)){
  L_Exp.post2015[,i] <-  Linf[i] * (1.0 - exp(-k[i] * (A[i] + time.vec)))
  time.mat.post2015[,i] <- A[i] + time.vec
}

colnames(L_Exp.post2015) <- unique.ID.post2015
L_Exp_long.post2015 <- melt(L_Exp.post2015)
colnames(L_Exp_long.post2015) <- c("time1", "ID", "CCL")

colnames(time.mat.post2015) <- unique.ID.post2015
time_long.post2015 <- melt(time.mat.post2015)
colnames(time_long.post2015) <- c("time1", "ID", "Years")

L_Exp_long.post2015$Years <- time_long.post2015$Years

ggplot(data = L_Exp_long.post2015) + 
  geom_path(aes(x = Years, y = CCL, color = ID)) +
  theme(legend.position = "none")
```


Compare k parameters between the time periods:

```{r}

k.pre2015 <- extract.posterior("k", jm.3$samples)
k.pre2015.stats <- data.frame(q50 = apply(k.pre2015, 
                                          MARGIN = 2, 
                                          FUN = median),
                              q2.5 = apply(k.pre2015,
                                           MARGIN = 2,
                                           FUN = quantile, 0.025),
                              q97.5 = apply(k.pre2015,
                                            MARGIN = 2,
                                            FUN = quantile, 0.975))
summary(k.pre2015.stats)

k.post2015 <- extract.posterior("k", jm.4$samples)
k.post2015.stats <- data.frame(q50 = apply(k.post2015, 
                                           MARGIN = 2, 
                                           FUN = median),
                               q2.5 = apply(k.post2015,
                                            MARGIN = 2,
                                            FUN = quantile, 0.025),
                               q97.5 = apply(k.post2015,
                                             MARGIN = 2,
                                             FUN = quantile, 0.975))
summary(k.post2015.stats)
```

```{r}
ggplot() + 
  geom_histogram(data = k.pre2015.stats,
                 aes(x = q50),
                 fill = "green",
                 binwidth = 0.02) + 
  geom_histogram(data = k.post2015.stats,
                 aes(x = q50),
                 fill = "blue",
                 binwidth = 0.02)
```


```{r}
ggplot() + 
  geom_histogram(data = k.pre2015.stats,
                 aes(x = q97.5),
                 fill = "green",
                 binwidth = 0.03) + 
  geom_histogram(data = k.post2015.stats,
                 aes(x = q97.5),
                 fill = "blue",
                 binwidth = 0.03)
```

```{r}
ggplot() + 
  geom_histogram(data = k.pre2015.stats,
                 aes(x = q2.5),
                 fill = "green",
                 binwidth = 0.008) + 
  geom_histogram(data = k.post2015.stats,
                 aes(x = q2.5),
                 fill = "blue",
                 binwidth = 0.008)
```

